A fast and Low-Overhead Technique to Secure Programs Against Integer Overflows

### 2012-09-07 doc/Manuscript/CGO/paper.tex

%Analysis about the amount of instructions
\begin{figure}[htb]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|}
\hline
Benchmark & \#I & \#II & \#II/\#I & \#O \\ \hline
470.lbm & 13,724 & 1,142 & 8.32\% & 0 \\ \hline
433.milc & 44,236 & 1,602 & 3.62\% & 11 \\ \hline
444.namd & 100,276 & 3,234 & 3.23\% & 12 \\ \hline
447.dealII & 1,381,408 & 36,157 & 2.62\% & 50 \\ \hline
450.soplex & 136,367 & 3,158 & 2.32\% & 13 \\ \hline
464.h264ref & 271,627 & 13,846 & 5.10\% & 167 \\ \hline
473.astar & 19,243 & 857 & 4.45\% & 0 \\ \hline
458.sjeng & 54,051 & 2,504 & 4.63\% & 68 \\ \hline
429.mcf & 4,725 & 165 & 3,49\% & 8 \\ \hline
471.omnetpp & 203,201 & 1,972 & 0.97\% & 2 \\ \hline
403.gcc & 1,419,456 & 18,669 & 1.32\% & N/A \\ \hline
445.gobmk & 308,475 & 14,129 & 4.58\% & 4 \\ \hline
462.libquantum & 16,297 & 928 & 5.69\% & 7 \\ \hline
401.bzip2 & 38,831 & 2,158 & 5.56\% & N/A \\ \hline
456.hmmer & 114,136 & 4,001 & 3.51\% & 0 \\ \hline
\multicolumn{1}{|r|}{Total (Average)} & 275.070 & 6.968 & 3,96\% &  \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:unprunedTable}
Instrumentation library without support of static range analysis.
\#I: number of LLVM bitcode instructions in the original program.
\#II: number of instructions that have been instrumented.
\#O: number of actual overflow tests detected dynamically.}
\end{figure}

@@

%Classification analysis
\begin{figure}[t!]
\begin{center}
\begin{scriptsize}
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
\hline
Bench & \#Sf & \#S & \#U & \#SO & \#SO/\#O & \#UO & \#UO/\#O
\\ \hline
lbm     & 1138 & 0    & 4 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
milc    & 536  & 17   & 1048  & 0 & 0,00\% & 11 & 1,05\% \\ \hline
namd    & 334  & 480  & 2420  & 0 & 0,00\% & 12 & 0,50\% \\ \hline
dealII  & 6188 & 39   & 28740  & 0 & 0,00\% & 50 & 0,17\% \\ \hline
soplex  & 229  & 16   & 2881 & 0 & 0,00\% & 13 & 0,45\% \\ \hline
h264ref & 2539 & 1195 & 10147 & 7 & 0,59\% & 167 & 1,65\% \\ \hline
astar   & 48   & 11   & 795 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
sjeng   & 150  & 213  & 1977 &0 & 0,00\% & 68 & 3,44\% \\ \hline
mcf     & 1    & 0    & 164 & 0 & 0,00\% & 8 & 4,88\% \\ \hline
omnetpp & 659  & 25   & 1288 & 1 & 4,00\% & 2 & 0,16\% \\ \hline
gcc     & 3365 & 1045 & 14065 & N/A & N/A & N/A & N/A \\ \hline
gobmk   & 1509 & 742  & 11736 & 0 & 0,00\% & 4 & 0,03\% \\ \hline
libqtum & 104  & 12   & 805 & 0 & 0,00\% & 7 & 0,87\% \\ \hline
bzip2   & 192  & 40   & 1926 & N/A & N/A & N/A & N/A \\ \hline
hmmer   & 663  & 222  & 3082 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
\end{tabular}
\end{scriptsize}
\end{center}
\caption{\label{fig:classificationTable}
How the range analysis classified arithmetic instructions.
\#Sf: safe.
\#S: suspicious.
\#U: uncertain.
\#SO: number of suspicious instructions that overflowed.
\#UO: number of uncertain instructions that overflowed.
}
\end{figure}

@@

Figure~\ref{fig:spec_runtime} shows the impact of our instrumentation in the
runtime of the SPEC benchmarks.
We run each benchmark 20 times.
The largest slowdown that we have observed, 11.83\%, happened in
\texttt{h264ref}, the benchmark that presented the largest number of distinct
sites where overflows happened dynamically. 
On the average, the instrumented programs are 3.24\% slower than the
original benchmarks.
If we use the range analysis to eliminate overflow checks, this slowdown falls
to 1.73\%.
The range analysis, in this case, reduces the instrumentation overhead by
53.40\%.
This improvement is larger than the percentage of overflow checks that we
avoid, e.g,. 20.57\%.
We believe that this difference is due to the fact that we are able to
eliminate checks on induction variables, as our range analysis can rely on
the loop boundaries to achieve this end.

@@

\noindent
\textbf{Dynamic Detection of Integer Overflows: }
Brumley {\em et al.}~\cite{Brumley07} have developed a tool, RICH, to
dynamically detect integer overflows.
Their goal is to secure C programs against integer overflow based
vulnerabilities.
The author's approach consists in instrumenting every integer operation that
might cause an overflow, underflow, or data loss.
The main result of Brumley {\em et al.} is the verification that instrumenting
programs to guard them against integer overflows does not compromise their
performance significantly: the average slowdown across four large applications
is 5\%.
RICH, Brumley {\em et al}'s tool, uses specific features of the x86 architecture
to reduce the instrumentation overhead.
Chinchani {\em et al.}~\cite{Chinchani04} follow a similar approach.
In this work, the authors describe each arithmetic operation formally, and then
use characteristics of the computer architecture to detect overflows at
runtime.
% TODO bibtex: cite Archerr: Runtime environment driven program safety.
Contrary to these previous work, we instrument programs at LLVM's intermediate
representation level, which is machine independent.
Nevertheless, the performance of the programs that we instrument is on par with
Brumley's, even without the support of the static range analysis.

@@

\noindent
\textbf{Static Detection of Integer Overflows: }
Zhang {\em et al.}~\cite{Zhang09} have used static analysis to sanitize
programs against integer overflow based vulnerabilities.
They instrument integer operations in paths from a source to a sink.
In Zhang {\em et al.}'s context, sources are functions that read values from
users, and sinks are memory allocation operations.
Thus, contrary to our work, Zhang {\em et al.}'s only need to instrument
about 10\% of the integer operations in the program.
However, they do not use any form of range analysis to limit the number of
checks inserted in the transformed code.
Wang {\em et a.}~\cite{Wang09} have implemented a tool, IntScope, that combines symbolic execution and taint analysis to detect integer overflow vulnerabilities. The authors have been able to use this tool to successfully identify many vulnerabilities in industrial quality software. Our work, and Wang {\em et a.}'s work are essentially different: they use symbolic execution, whereas we rely on
range analysis.
Furthermore, contrary to them, we do not try to detected statically the
possibility of an integer overflow happening.
Contrary to us, they do not transform the program to prevent or detect such event dynamically.
Still in the field of symbolic execution, Molnar
{\em et al.}~\cite{Molnar09} have implemented a tool, SmartFuzz, that analyzes
Linux x86 binaries to find integer overflow bugs.
They prove the existence of bugs by generating test cases for them.

>>>>

%Analysis about the amount of instructions
\begin{figure}[htb]
\begin{center}
\begin{small}
\begin{tabular*}{\columnwidth}{@{\extracolsep{\fill}}|l|r|r|r|r|}
\hline
Benchmark & \#I & \#II & \#II/\#I & \#O \\ \hline
470.lbm & 13,724 & 1,142 & 8.32\% & 0 \\ \hline
433.milc & 44,236 & 1,602 & 3.62\% & 11 \\ \hline
444.namd & 100,276 & 3,234 & 3.23\% & 12 \\ \hline
447.dealII & 1,381,408 & 36,157 & 2.62\% & 50 \\ \hline
450.soplex & 136,367 & 3,158 & 2.32\% & 13 \\ \hline
464.h264ref & 271,627 & 13,846 & 5.10\% & 167 \\ \hline
473.astar & 19,243 & 857 & 4.45\% & 0 \\ \hline
458.sjeng & 54,051 & 2,504 & 4.63\% & 68 \\ \hline
429.mcf & 4,725 & 165 & 3,49\% & 8 \\ \hline
471.omnetpp & 203,201 & 1,972 & 0.97\% & 2 \\ \hline
403.gcc & 1,419,456 & 18,669 & 1.32\% & N/A \\ \hline
445.gobmk & 308,475 & 14,129 & 4.58\% & 4 \\ \hline
462.libquantum & 16,297 & 928 & 5.69\% & 7 \\ \hline
401.bzip2 & 38,831 & 2,158 & 5.56\% & N/A \\ \hline
456.hmmer & 114,136 & 4,001 & 3.51\% & 0 \\ \hline
\multicolumn{1}{|r|}{Total (Average)} & 275.070 & 6.968 & 3,96\% &  \\ \hline
\end{tabular*}
\end{small}
\end{center}
\caption{\label{fig:unprunedTable}
Instrumentation library without support of static range analysis.
\#I: number of LLVM bitcode instructions in the original program.
\#II: number of instructions that have been instrumented.
\#O: number of instructions that actually overflowed in the dynamic tests.}
\end{figure}

@@

%Classification analysis
\begin{figure}[t!]
\begin{center}
\begin{scriptsize}
\begin{tabular}{|l|r|r|r|r|r|r|r|r|}
\hline
Bench & \#Sf & \#S & \#U & \#SO & \#SO/\#S & \#UO & \#UO/\#U
\\ \hline
lbm     & 1138 & 0    & 4 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
milc    & 536  & 17   & 1048  & 0 & 0,00\% & 11 & 1,05\% \\ \hline
namd    & 334  & 480  & 2420  & 0 & 0,00\% & 12 & 0,50\% \\ \hline
dealII  & 6188 & 39   & 28740  & 0 & 0,00\% & 50 & 0,17\% \\ \hline
soplex  & 229  & 16   & 2881 & 0 & 0,00\% & 13 & 0,45\% \\ \hline
h264ref & 2539 & 1195 & 10147 & 7 & 0,59\% & 160 & 1,58\% \\ \hline
astar   & 48   & 11   & 795 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
sjeng   & 150  & 213  & 1977 &0 & 0,00\% & 68 & 3,44\% \\ \hline
mcf     & 1    & 0    & 164 & 0 & 0,00\% & 8 & 4,88\% \\ \hline
omnetpp & 659  & 25   & 1288 & 1 & 4,00\% & 1 & 0,07\% \\ \hline
gcc     & 3365 & 1045 & 14065 & N/A & N/A & N/A & N/A \\ \hline
gobmk   & 1509 & 742  & 11736 & 0 & 0,00\% & 4 & 0,03\% \\ \hline
libqtum & 104  & 12   & 805 & 0 & 0,00\% & 7 & 0,87\% \\ \hline
bzip2   & 192  & 40   & 1926 & N/A & N/A & N/A & N/A \\ \hline
hmmer   & 663  & 222  & 3082 & 0 & 0,00\% & 0 & 0,00\% \\ \hline
\end{tabular}
\end{scriptsize}
\end{center}
\caption{\label{fig:classificationTable}
How the range analysis classified arithmetic instructions in the 
u-SSA form program.
\#Sf: safe.
\#S: suspicious.
\#U: uncertain.
\#SO: number of suspicious instructions that overflowed.
\#UO: number of uncertain instructions that overflowed.
}
\end{figure}

@@

Figure~\ref{fig:spec_runtime} shows the impact of our instrumentation in the
runtime of the SPEC benchmarks.
We run each benchmark 20 times.
The largest slowdown that we have observed, 11.83\%, happened in
\texttt{h264ref}, the benchmark that presented the largest number of distinct
sites where overflows happened dynamically. 
On the average, the instrumented programs are 3.24\% slower than the
original benchmarks.
If we use the range analysis to eliminate overflow checks, this slowdown falls
to 1.73\%.
The range analysis, in this case, reduces the instrumentation overhead by
46.60\%.
This improvement is larger than the percentage of overflow checks that we
avoid, e.g,. 20.57\%.
We believe that this difference is due to the fact that we are able to
eliminate checks on induction variables, as our range analysis can rely on
the loop boundaries to achieve this end.

@@

\noindent
\textbf{Dynamic Detection of Integer Overflows: }
Brumley {\em et al.}~\cite{Brumley07} have developed a tool, RICH, to
secure C programs against integer overflow based vulnerabilities.
The author's approach consists in instrumenting every integer operation that
might cause an overflow, underflow, or data loss.
The main result of Brumley {\em et al.} is the verification that instrumenting
programs to guard them against integer overflows does not compromise their
performance significantly: the average slowdown across four large applications
is 5\%.
RICH, Brumley {\em et al}'s tool, uses specific features of the x86 architecture
to reduce the instrumentation overhead.
Chinchani {\em et al.}~\cite{Chinchani04} follow a similar approach.
In this work, the authors describe each arithmetic operation formally, and then
use characteristics of the computer architecture to detect overflows at
runtime.
% TODO bibtex: cite Archerr: Runtime environment driven program safety.
Contrary to these previous work, we instrument programs at LLVM's intermediate
representation level, which is machine independent.
Nevertheless, the performance of the programs that we instrument is on par with
Brumley's, even without the support of the static range analysis.

@@

\noindent
\textbf{Static Detection of Integer Overflows: }
Zhang {\em et al.}~\cite{Zhang09} have used static analysis to sanitize
programs against integer overflow based vulnerabilities.
They instrument integer operations in paths from a source to a sink.
In Zhang {\em et al.}'s context, sources are functions that read values from
users, and sinks are memory allocation operations.
Thus, contrary to our work, Zhang {\em et al.}'s only need to instrument
about 10\% of the integer operations in the program.
However, they do not use any form of range analysis to limit the number of
checks inserted in the transformed code.
Wang {\em et a.}~\cite{Wang09} have implemented a tool, IntScope, that combines symbolic execution and taint analysis to detect integer overflow vulnerabilities. The authors have been able to use this tool to successfully identify many vulnerabilities in industrial quality software. Our work, and Wang {\em et a.}'s work are essentially different: they use symbolic execution, whereas we rely on
range analysis.
Contrary to us, they do not transform the program to prevent or detect such event dynamically.
Still in the field of symbolic execution, Molnar
{\em et al.}~\cite{Molnar09} have implemented a tool, SmartFuzz, that analyzes
Linux x86 binaries to find integer overflow bugs.
They prove the existence of bugs by generating test cases for them.
